Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-03-16 paper/zhang-cook-laa-langrene-menendez.Rmd
fix various plots and writing on interestingness of the index

Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions. A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix. We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection. In projection pursuit, higher index values are associated with interesting projections that depart from normality [@diaconis1984asymptotics; @huber1985projection], e.g. Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial] and natural Hermite index [@cook1993projection] for measuring the distance between the density of projected data and a standard normal distribution, LDA index [@lee2005projection] and PDA index [@lee2010projection] for finding clustering structure in the data, and skewness-based index [@Loperfido2018] and kurtosis-based index [@Loperfido2020] for outlier detection.

@@

```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index. Optimiser CRS is used before the polish. The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis."}
```

@@

Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by CRS and followed by the polish. The projection by the final basis of each algorithm is also shown. The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut. Polish works with this end basis and further pushes the index value to produce much clear edges of the cluster, especially along the vertical axis.

@@

```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}. Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index. Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best."}
```

>>>>

Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions. A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix. We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value. This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection. An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal as $n$ and $p$ approaching infinity under certain conditions. @huber1985projection illustrates that, for indices based on entropy measures, [...]. Index functions  proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution.(Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]); chi-square index [@posse1995projection] for detecting spiral structure; kurtosis [@Loperfido2020] and skewness [@Loperfido2018] indices for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.

@@

```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index. Top row shows the intial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertial axis."}
```

@@

Figure \ref{fig:polish} presents the projectd data and trace plot of a 2D projection, optimised by CRS and followed by the polish. In the top row shows the intial projection, the final projection after CRS, and the final projection after polish, repsectively. The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut. Polish works with this end basis and further pushes the index value to produce much clear edges of the cluster, especially along the vertical axis.

@@

```{r noisy-better-geo, fig.height = 5, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}. Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index. Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best."}
```

### 2021-03-16 paper/zhang-cook-laa-langrene-menendez.tex
fix various plots and writing on interestingness of the index

Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
observations in \(p\) dimensions. A \(d\)-dimensional projection is a
linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{Y}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis matrix. We define
\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projected data \(\mathbf{Y}\) onto a scalar
value. This is commonly known as the projection pursuit index function,
or just index function, and is used to measure the ``interestingness''
of a given projection. In projection pursuit, higher index values are
associated with interesting projections that depart from normality
\citep{diaconis1984asymptotics, huber1985projection}, e.g.~Legendre
index \citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial} and natural Hermite index
\citep{cook1993projection} for measuring the distance between the
density of projected data and a standard normal distribution, LDA index
\citep{lee2005projection} and PDA index \citep{lee2010projection} for
finding clustering structure in the data, and skewness-based index
\citep{Loperfido2018} and kurtosis-based index \citep{Loperfido2020} for
outlier detection.

@@

\caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index. Optimiser CRS is used before the polish. The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis.}\label{fig:polish}
\end{figure}
\end{Schunk}

>>>>

Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
observations in \(p\) dimensions. A \(d\)-dimensional projection is a
linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{Y}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis matrix. We define
\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projected data \(\mathbf{Y}\) onto a scalar
value. This is commonly known as the projection pursuit index function,
or just index function, and is used to measure the ``interestingness''
of a given projection. An interesting projection shows structures that
are non-normal and this has been argued for two reasons: Theoretical
proofs from \citet{diaconis1984asymptotics} show that projections tend
to be normal as \(n\) and \(p\) approaching infinity under certain
conditions. \citet{huber1985projection} illustrates that, for indices
based on entropy measures, {[}\ldots{}{]}. Index functions proposed in
the litearture include early indices that can be categorised by
measuring the \(L^2\) distance between the projection and a normal
distribution.(Legendre index \citep{friedman1974projection}, Hermite
index \citep{hall1989polynomial}, and natural Hermite index
\citep{cook1993projection}); chi-square index
\citep{posse1995projection} for detecting spiral structure; kurtosis
\citep{Loperfido2020} and skewness \citep{Loperfido2018} indices for
detecting outliers in financial time series; and most recently,
scagnostic indices \citep{laa2020using} for summarising structures in
scatterplot matrices based on eight scagnostic measures.

@@

\caption[Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index. Top row shows the intial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertial axis.}\label{fig:polish}
\end{figure}
\end{Schunk}
